Appendix: Why the complement of AUC isn't a casewise CV criterion

### 2023-12-30 scratch/notes_georges/AUC_not_casewise-edited-1.Rmd
tweak AUC appendix

The following R function computes AUC from $\mathbf{\widehat{y}}$ and $\mathbf{y}$:

```{r}
AUC <- function(y, yhat = seq_along(y)) {
  # AUC:
  # y is a vector of 0s and 1s
  # yhat is a predictor of y, only used for its order
  #    i.e. the result is invariant under monotonic transformations of yhat
  #    If there are ties in yhat, this function splits the ties
  s <- sum(y)
  n_v <- length(y)
  if (s == 0) return(0)
  if (s == n_v) return(1)
  a <- sum(outer(yhat[y == 0], yhat[y == 1], `<`)) + 
    sum(outer(yhat[y == 0], yhat[y == 1], `==`))/2
  a/(s*(n_v - s))
}
```

@@

We can test whether equation (\ref{eq:lin}) has a solution for any given $n_v$ by considering whether the residuals of the associated linear model are all 0, using the "design matrix" $[\mathbf{U} -\mathbf{Y}, \mathbf{Y}]$ to predict the outcome
$\mathrm{AUC}(\mathbf{\widehat{Y}},\mathbf{Y})$.

>>>>

The following R function computes AUC from $\mathbf{\widehat{y}}$ and $\mathbf{y}$, accommodating the cases where $\mathbf{y}$ is all 0s or all 1s:

```{r}
AUC <- function(y, yhat = seq_along(y)) {
  s <- sum(y)
  if (s == 0) return(0)
  if (s == length(y)) return(1)
  Metrics::auc(y, yhat)
}
```

@@

We can test whether equation (\ref{eq:lin}) has a solution for any given $n_v$ by trying to solve it as a least-squares problem, considering whether the residuals of the associated linear model are all 0, using the "design matrix" $[\mathbf{U} -\mathbf{Y}, \mathbf{Y}]$ to predict the "outcome"
$\mathrm{AUC}(\mathbf{\widehat{Y}},\mathbf{Y})$:
